---
title: "A498 and 786-O Differential gRNA abundance analysis"
author: "Jeremy M. Simon, Xiaosai Yao"
date: "2/17/2023"
output:
  html_document: default
  pdf_document: default
---

# Load libraries, set working directory
```{r, message=FALSE, warning=FALSE}
library(DESeq2)
library(ComplexHeatmap)
library(paletteer)
library(eulerr)
```

```{r,echo=F}
setwd("/proj/jmsimon/Zhang/ccRCC_TCGA_GDC_dCas9screen")
```

Differential analysis of dCas9_KRAB + gRNA targeting top 500 gained enhancer regions in A498 and 786O

Top enhancers based on Yao, Cancer Discovery 2017

Experiment performed by Tyler Klann of Charles Gersbach lab of Duke University



# Define differential analysis pvalue and log fold change cutoff
Note that pvalue is used in place of p-adjusted to allow more exclusive secondary screening later on
```{r}
pcutoff <- 0.05
logFCcutoff <- -1.5
```

# Perform differential analysis for A498
```{r}
# Load previously saved results
batch_a498 <- read.table('data/batch_a498.txt', header=TRUE, sep='\t',row.names=1,stringsAsFactors=FALSE)

A498_matrix <- read.table('data/A498-krab-gained-enh.txt',
                          header=TRUE, 
                          sep='\t',
                          row.names=1,
                          stringsAsFactors=FALSE)


A498_dds <- DESeqDataSetFromMatrix(countData = A498_matrix,
                                   colData = batch_a498,
                                   design = ~ condition)


# Run differential abundance test
A498_dds <- DESeq(A498_dds, fitType="local")
A498_dds <- estimateSizeFactors(A498_dds)
A498_par <- estimateDispersions(A498_dds, fitType = "parametric")
A498_loc <- estimateDispersions(A498_dds, fitType = "local")
plotDispEsts(A498_par, main= "dispEst: parametric")

plotDispEsts(A498_loc, main= "dispEst: local")

A498_par_residual <- median(abs(mcols(A498_par)$dispGeneEst - mcols(A498_par)$dispFit))
A498_loc_residual <- median(abs(mcols(A498_loc)$dispGeneEst - mcols(A498_loc)$dispFit))

# Report results
A498_res <- results(A498_dds, contrast=c("condition","OneMonthCulture","Initial"))

length(which(A498_res$pvalue <0.05 & A498_res$log2FoldChange <0))

length(which(A498_res$pvalue <0.05 & A498_res$log2FoldChange >0))

length(which(A498_res$padj <0.1))
```

# Plot MA plot
In order to shade significant regions by pvalue, replace padj by pvalue just for plotting
```{r}
A498_res.MA <- A498_res
A498_res.MA$padj <- A498_res.MA$pvalue
plotMA(A498_res.MA, 
       alpha = 0.05 ,
       ylim=c(-5,5), 
       main = "A498", 
       colSig = "red")

```

# Plot PCA
```{r}
A498_vsd <- vst(A498_dds, blind=FALSE)
print(DESeq2::plotPCA(A498_vsd, intgroup=c("condition")))
```

# Plot cook's distance to check for outliers
```{r}
par(mar=c(8,5,2,2))
boxplot(log10(assays(A498_dds)[["cooks"]]), range=0, las=2)
```

# Plot dispersion
```{r}
plotDispEsts(A498_dds)
```

# Plot heatmap 
We will capture the output here for incorporation into summary figure later

```{r}
A498_rld <- rlog(A498_dds,fitType='local')
A498_rld.sig <- assay(A498_rld)[which(A498_res$pvalue < pcutoff & A498_res$log2FoldChange < logFCcutoff), ]
A498_rld.sig <- data.frame(A498_rld.sig)
A498_res$enhancerID <- rownames(A498_res)

# Median-center rows
medianCtr<-function(x){
    annAll <- dimnames(x)
    medians <- apply(x,1,median,na.rm=T)
    x <- t(scale(t(x),center=medians,scale=F))
    dimnames(x) <- annAll
    return(x)
}

A498_rld.sig.ctr <- medianCtr(A498_rld.sig)

ha1 <- HeatmapAnnotation(Condition = colData(A498_dds)$condition, 
	border=T, 
	col = list(Condition = c("Initial" = paletteer_d(`"colorblindr::OkabeIto"`)[2], "OneMonthCulture" = paletteer_d(`"colorblindr::OkabeIto"`)[1]))
)

plot1 <- ComplexHeatmap::Heatmap(na.omit(A498_rld.sig.ctr),
	cluster_columns=F,
	name = "Median-centered\nnormalized abundance",
	cluster_rows=F,
	show_row_names=F,
	border=T,
	column_title = "A498",
	row_title = paste0("n = ",nrow(na.omit(A498_rld.sig.ctr))),
	top_annotation = ha1,
	column_labels = c(1:3,1:3),
	col = circlize::colorRamp2(c(-2, 0, 2), c("blue","white","red"))
	)

draw(plot1, heatmap_legend_side="bottom", annotation_legend_side = "bottom")


saveRDS(plot1,'A498.heatmap.plot.rds')
write.table(A498_res,paste('Deseq2_A498_krab_R2.local.txt',sep=''),row.names=T, col.names=T,sep="\t")
```







# Perform differential analysis for 786-O
```{r}
# Load previously saved results
O786_matrix <- read.table('data/786O-krab-gained-enh.txt',
							header=T, 
							sep='\t',
							row.names=1,
							stringsAsFactors=F)

batch_786O <- read.table('data/batch_786O.txt', header=TRUE, sep='\t',row.names=1,stringsAsFactors=FALSE)

# Run differential abundance test
O786_dds <- DESeqDataSetFromMatrix(countData = O786_matrix,
                              colData = batch_786O,
                              design = ~ condition)

O786_dds <- estimateSizeFactors(O786_dds)
O786_par <- estimateDispersions(O786_dds, fitType = "parametric")
O786_loc <- estimateDispersions(O786_dds, fitType = "local")
plotDispEsts(O786_par, main= "dispEst: parametric")

plotDispEsts(O786_loc, main= "dispEst: local")

O786_par_residual <- median(abs(mcols(O786_par)$dispGeneEst - mcols(O786_par)$dispFit),na.rm = TRUE)
O786_loc_residual <- median(abs(mcols(O786_loc)$dispGeneEst - mcols(O786_loc)$dispFit),na.rm = TRUE)

# Choose local because residual was smaller

O786_dds <- DESeq(O786_dds, fitType="local")

# Report results
O786_res <- results(O786_dds, contrast=c("condition","OneMonthCulture","Initial"),cooksCutoff=FALSE)


length(which(O786_res$pvalue <0.05 & O786_res$log2FoldChange <0))
length(which(O786_res$pvalue <0.05 & O786_res$log2FoldChange >0))

length(which(O786_res$padj <0.1))
```

# Plot MA plot 
In order to shade by pvalue, replace padj by pvalue just for plotting
```{r}
O786_res.MA <- O786_res
O786_res.MA$padj <- O786_res.MA$pvalue

plotMA(O786_res.MA, 
       alpha=0.05, 
       main="786O", 
       ylim=c(-5,5), 
       colSig = "red")

```

# Plot PCA
```{r}
O786_vsd <- vst(O786_dds, blind=FALSE)
print(DESeq2::plotPCA(O786_vsd, intgroup=c("condition")))
```

# Plot cook's distance to check for outliers
```{r}
par(mar=c(8,5,2,2))
boxplot(log10(assays(O786_dds)[["cooks"]]), range=0, las=2)
```

# Plot dispersion
```{r}
plotDispEsts(O786_dds)
```


# Plot heatmap 
We will capture the output here for incorporation into summary figure later

```{r}
O786_rld <- rlog(O786_dds, fitType='local')
O786_rld.sig <- assay(O786_rld)[which(O786_res$pvalue < pcutoff & O786_res$log2FoldChange < logFCcutoff), ]
O786_rld.sig <- data.frame(O786_rld.sig)
O786_res$enhancerID <- rownames(O786_res)

# Median-center rows
O786_rld.sig.ctr <- medianCtr(O786_rld.sig)

ha1 <- HeatmapAnnotation(Condition = colData(O786_dds)$condition,
        border=T,
        col = list(Condition = c("Initial" = paletteer_d(`"colorblindr::OkabeIto"`)[2], "OneMonthCulture" = paletteer_d(`"colorblindr::OkabeIto"`)[1]))
)

plot2 <- ComplexHeatmap::Heatmap(na.omit(O786_rld.sig.ctr),
	cluster_columns=F,
        name = "Median-centered\nnormalized abundance",
	column_title = "786-O",
	row_title = paste0("n = ",nrow(na.omit(O786_rld.sig.ctr))),
        cluster_rows=F,
        show_row_names=F,
        border=T,
        top_annotation = ha1,
        column_labels = c(1:3,1:3),
	col = circlize::colorRamp2(c(-2, 0, 2), c("blue","white","red"))
)

draw(plot2, heatmap_legend_side="bottom", annotation_legend_side = "bottom")


saveRDS(plot2,'786O.heatmap.plot.rds')
write.table(O786_res,paste('Deseq2_786O_krab_local.txt',sep=''),row.names=TRUE, col.names=TRUE, sep="\t")
```


# Combine results
```{r}
colnames(O786_res) <- paste0("786O_",colnames(O786_res))
colnames(O786_matrix) <- paste0("786O_",colnames(O786_matrix))
colnames(A498_res) <- paste0("A498_",colnames(A498_res))
colnames(A498_matrix) <- paste0("A498_",colnames(A498_matrix))
both_res <- cbind(O786_res, O786_matrix[rownames(O786_res),], A498_res,  A498_matrix[rownames( A498_res),])
both_res <- data.frame(both_res)
head(both_res)

write.table(both_res,paste('Deseq2_786O_A498_krab_local.txt',sep=''),row.names=TRUE, col.names=TRUE,sep="\t")
```



# Save workspace
```{r}
save.image("dCas9screen_results.RData")
```

# Get session info
```{r}
sessionInfo()
```
