---
title: "A498 and 786-O Differential gRNA gene assignments"
author: "Jeremy M. Simon, Xiaosai Yao"
date: "2/20/2023"
output:
  html_document: default
  pdf_document: default
---

# Load libraries, set working directory
```{r, message=FALSE, warning=FALSE}
library(rGREAT)
library(tidyverse)
library(Vennerable)
```

```{r,echo=F}
setwd("/proj/jmsimon/Zhang/ccRCC_TCGA_GDC_dCas9screen")
```

# Load previously saved workspace
```{r}
load("dCas9screen_results.RData")
```


# Associate 786-O to genes using rGREAT linkages to two closest hg19 genes
In future could improve if updated version allows inclusion of 'name' field in input, see [here](https://github.com/jokergoo/rGREAT/issues/38). 
For now merging based on coordinates
```{r}

O786_sig_regions_tbl <- enframe(rownames(O786_rld.sig),name=NULL) %>% 
    as_tibble() %>%
    tidyr::extract(value,c("chr","start","end","name"),"(^.+)_(.+)_(.+)_(enh.+)",remove=F) %>%
	dplyr::select(chr,start,end,value) %>%
	dplyr::rename("name" = value) %>%
	mutate(start = as.numeric(start), end = as.numeric(end))

O786_GREATjob <- submitGreatJob(as.data.frame(O786_sig_regions_tbl),
	species="hg19",
	rule="twoClosest",
	gr_is_zero_based=T)

# Peek at gene associations
getRegionGeneAssociations(O786_GREATjob)

O786_sig_merged <- as.data.frame(getRegionGeneAssociations(O786_GREATjob)) %>% 
	as_tibble() %>% 
	unnest_wider(c(annotated_genes,dist_to_TSS),names_sep="_") %>%
	dplyr::rename("chr" = seqnames) %>%
	dplyr::select(-width,-strand) %>%
	inner_join(O786_sig_regions_tbl,by=c("chr","start","end")) %>%
	inner_join(as.data.frame(O786_res) %>% as_tibble(),by=c("name" = "X786O_enhancerID")) %>%
	dplyr::select(chr, 
		start, 
		end, 
		name, 
		baseMean = X786O_baseMean, 
		log2FoldChange = X786O_log2FoldChange,
		lfcSE = X786O_lfcSE,
		stat = X786O_stat,
		pvalue = X786O_pvalue,
		padj = X786O_padj,
		annotated_genes_1,
		annotated_genes_2,
		dist_to_TSS_1,
		dist_to_TSS_2)

write_tsv(O786_sig_merged,"786-O_dCas9_sig_withLinkedGenes.tsv")

O786_sig_merged_genes <- O786_sig_merged %>%
	dplyr::select(contains("annotated_genes")) %>%
	unclass() %>% 
	stack() %>%
	pull(values) %>%
	unique() %>%
	.[!is.na(.)]
```

```{r,echo=F}
Sys.sleep(60)
```

# Associate A498 to genes using rGREAT linkages to two closest hg19 genes
In future could improve if updated version allows inclusion of 'name' field in input, see [here](https://github.com/jokergoo/rGREAT/issues/38).
For now merging based on coordinates
```{r}

A498_sig_regions_tbl <- enframe(rownames(A498_rld.sig),name=NULL) %>% 
    as_tibble() %>%
    tidyr::extract(value,c("chr","start","end","name"),"(^.+)_(.+)_(.+)_(enh.+)",remove=F) %>%
	dplyr::select(chr,start,end,value) %>%
	dplyr::rename("name" = value) %>%
	mutate(start = as.numeric(start), end = as.numeric(end))

A498_GREATjob <- submitGreatJob(as.data.frame(A498_sig_regions_tbl),
	species="hg19",
	rule="twoClosest",
	gr_is_zero_based=T)

# Peek at gene associations
getRegionGeneAssociations(A498_GREATjob)

A498_sig_merged <- as.data.frame(getRegionGeneAssociations(A498_GREATjob)) %>% 
	as_tibble() %>% 
	unnest_wider(c(annotated_genes,dist_to_TSS),names_sep="_") %>%
	dplyr::rename("chr" = seqnames) %>%
	dplyr::select(-width,-strand) %>%
	inner_join(A498_sig_regions_tbl,by=c("chr","start","end")) %>%
	inner_join(as.data.frame(A498_res) %>% as_tibble(),by=c("name" = "A498_enhancerID")) %>%
	dplyr::select(chr, 
		start, 
		end, 
		name, 
		baseMean = A498_baseMean, 
		log2FoldChange = A498_log2FoldChange,
		lfcSE = A498_lfcSE,
		stat = A498_stat,
		pvalue = A498_pvalue,
		padj = A498_padj,
		annotated_genes_1,
		annotated_genes_2,
		dist_to_TSS_1,
		dist_to_TSS_2)

write_tsv(A498_sig_merged,"A498_dCas9_sig_withLinkedGenes.tsv")

A498_sig_merged_genes <- A498_sig_merged %>%
	dplyr::select(contains("annotated_genes")) %>%
	unclass() %>% 
	stack() %>%
	pull(values) %>%
	unique() %>%
	.[!is.na(.)]
```


# Count number of intersecting linked genes and plot Venn diagram
```{r}
length(intersect(O786_sig_merged_genes,A498_sig_merged_genes))

vlist <- list("786-O" = O786_sig_merged_genes, "A498" = A498_sig_merged_genes)
v <- Venn(vlist)
plot(v)

# Print list of genes
overlapping_genes <- intersect(O786_sig_merged_genes,A498_sig_merged_genes)
enframe(overlapping_genes,name=NULL) %>%
	dplyr::arrange(value) %>%
	print(n=100)

```

# Save workspace
```{r}
save.image("dCas9screen_assignToGenes.RData")
```

# Get session info
```{r}
sessionInfo()
```
